\documentclass{article}
\usepackage[top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}
\usepackage{hyperref}

% Add periods after title numbers (chapters and sections and in table of contents)
\usepackage{titlesec}
\titleformat{\section}{\bfseries\large}{\thesection}{.5em}{}[\hrule]
\renewcommand{\thesection}{\thechapter\arabic{section}.}
\renewcommand{\thesubsection}{\thesection\arabic{subsection}.}

% As we won't use paragraph{} statement, we are gonna have to redefine the
% inter-line space.
\usepackage{setspace}
%\singlespacing
\onehalfspacing

% Indent the beginning of every section
\usepackage{indentfirst}

% Ask to have space between 2 paragraphs.
\usepackage{parskip}

\usepackage{listings}
\lstset{
    basicstyle=\normalsize\ttfamily,
    frame=single,
    columns=flexible,
    breaklines=true,
    % Used because the Vim LaTeX colorscheme is buggy, using a $ character will
    % break the colorscheme. Defining a <dollar> statement avoids to use
    % a $ character.
    literate={<dollar>}{\$}1
}

\begin{document}

\title{\underline{L}ibre\underline{O}ffice \underline{W}eekly \underline{N}ews}
\author{William Gathoye}
\date{June 14, 2014}
\maketitle



\section{\texorpdfstring{Welcome to this first edition of
\underline{L}ibre\underline{O}ffice \underline{W}eekly
\underline{N}ews!}{}}

LibreOffice has plenty of communication channels for its developers and
contributors to exchange messages: mailing lists, social networks, wiki and an
ask (question/answers) instance. This configuration is not ideal for the
developer to see a global overview of the progress made in the LibreOffice
development. Since I subscribed most of the LibreOffice mailing lists and
usually take the time to read all messages I receive in my mailbox, and after
a request made by Michal Meeks, I decided to take up this challenge.

Like \emph{Linux Weekly News} at its early life, LibreOffice Weekly News
consists of a weekly issue gathering interesting development reported in
discussions (mailing lists or IRC), Q/A and git commits made over the past
week. This summary will emphase on the developer point of view; so do not
expect to have a popularized article, and please do not complain about nerdy
stuff (if any), since it is written by a computer science student (me) for
developers. 

I will publish this summary on the LibreOffice developer mailing list each week
on Sunday nights. In that way, developers not working on LibreOffice during the
week-end can even get the evolution made throughout the weekend. A beta version
of the article will be made available on my github account\cite{wgetGithub} on
Fridays evening. In that way, if someone wants to proofread this writing and
apply corrections before the final version released on Sundays, he is more than
welcome.

This week in LibreOffice\ldots



\section{GUI.}

The conversion towards the .ui glade dialogs is now continuing (thanks Szymon
Kłois!) \cite{gladeUi1,gladeUi2,gladeUi3}.

Seems like the new start center is still considered as "unfinished" regarding
the request we have from users. These one are asking for an option to disable
document thumbnail on the new start
screen\cite{startCenterHidePreview1}\cite{startCenterHidePreview2}. Maybe this
should be considered as an easy hack?

As now, we have still no news from the team responsible for the (great!) Sifr
icon theme. The problem for now is that we have 2 inconsistent repositories for
Sifr: \begin{itemize} \item the freedesktop official where all the libreoffice
            code is hosted \item and a Github account created to help people
                accustomed to use Github, but scared about freedesktop, to get
                involved more easily.  \end{itemize}

The initial idea was to merge the Github account into freedesktop. According to
an answer late in this week\cite{sifrGithub1}, icon designers are using
SparkleShare with Github. SparkleShare is an application acting like DropBox
but with a commit each time a file is pushed or modified. The GitHub account
has thus uncompleted work, because it is used like a drive in the cloud. We
have to handle this issue, by creating 2 branches, keep both repos in sync and
give designers commit right only to the sifr folder on freedesktop.



\section{Security.}

OpenSSL patches from Fedora have been committed\cite{opensslPatches} (Caolán
McNamara) to the libreoffice-4-2 branch. As future branches
(\lstinline{libreoffice-4-3} and \lstinline{master}) are using newer and
unaffected OpenSSL version, there was no need to include them.

\section{Code base evolution.}

Since Chrom(e$\vert$ium) dropped the old NPAPI and Firefox is moving away from
browser extensions\cite{npapiMozilla}, it was high time for LibreOffice to
remove that old code\cite{removeNpapiPatch} (thanks Bryan Quigley), not working
in some scenarios\cite{npapiLibONotWorking}.

There is still a bunch of features that are copied/implemented at different
locations in the LibreOffice code base. We saw some progress on a copy/paste
code detector tool\cite{copyPasteDetectorBug}. The tool "CCFinderX core"
available on github under MIT license could suit LibreOffice
nicelly\cite{copyPasteDetectorGithub}.

Access2Base (see \cite{access2BaseDefinition} for a definition) is now included
in libreoffice 4.2 and later. In order to keep extension compatibility with
other LibreOffice/OO.o forks, Access2Base will be provided as an extension to
them. It was not possible to provide Access2Base as an extension for
LibreOffice 4.2 and latter because, LibreOffice is refusing to override
internal parts of LibreOffice (this still generates discussions with great
argumentation weighing pros and cons\cite{coreOverrideArguments}). Thanks to
this patch\cite{access2BaseOverridePatch} (Lionel Elie Mamane), it is now
possible for extensions to override Access2Base.



\section{Build environment.}

This week, we have seen an initial discussion about reducing build time on
Windows by using more native tools.

To be built on Windows, LibreOffice still uses Cygwin. As Cygwin provides
a rather great package management, installing dependencies on Windows is quite
easy [9]. So dropping Cygwin in favor of smaller/native tools like GnuWin32 or
busybox is out of scope for now (otherwize this would require the dev of
a one-click tool to install the needed
dependencies\cite{winBuildDependenciesTool} and maybe reopen this
bug\cite{winBuildDependenciesInstall}.

Providing upstream patches to 'make' came up during the discussion.

But now, the discussion seems limited to this question: are we gonna replace
gmake from a Cygwin-running version to a native-version provided by GnuWin32
which should provide speedup improvement over Cygwin. If it works on Windows
tinderbuilds, this could be added as an autodetection option to autogen:
defining MAKE or a GNUMAKE environment variable manually could do the trick. In
this case there is no need for
a synchronization/d-day\cite{winBuildGmakeReplaceGnuWin32}.

Cross compilation on Linux with MinGW is rather annoying since it break
compatibility with ABI and extensions used on Cygwin/VisualC++ builds. And we
can see how this compatibility problem of extensions could be
critical\cite{extensionsCompatibility}.

For a reminder about the differences between Cygwin, GNUWin32 and MSYS, you can
read this\cite{winBuildCygwinMsysGnuWin32Comparison}.

Sometimes we have bug appearing in some languages while most of others have no
problem. Since localization takes much time, it has been disabled on
tinderboxes. After a request, thinderbox nb 51 compiling for Windows will now
produce builds for \lstinline{libreoffice-4-3} and \lstinline{master} branches
with the following languages:
\begin{itemize}
    \item de (as one language with long labels);
    \item ar (as RTL language);
    \item ja (as CJK representative); 
    \item and qtz (keyID-pseudo-language).  
\end{itemize}

LibreOffice is now buildable on FreeBSD
again\cite{freeBsdBuild1}\cite{freeBsdBuild2}.



\section{Unit tests.}

We now have the results (thanks Markus Mohrhard!) of a memcheck run against 20k
documents that can be used to find nasty problem in future
development\cite{memcheckTests}. Two problems have been fixed during these
operations: (according to a chat with mmeeks) 4x the current speed in Writer
when trying to open large documents.

\begin{lstlisting}
(2014-06-13 15:17:03) ***mmeeks has a 4x speedup for loading large writer files
he's working on finishing =)
\end{lstlisting}



\section{Other issues.}

The MySQL/MariaDB connector appears to be broken on OS X. Using homebrew is
discouraged. ODBC code assumes Unicode char is 2 bytes which is
wrong\cite{sqlConnectorOsx1}\cite{sqlConnectorOsx2}.

An issue stating that Writer doesn't memorize the default printer defined in
Windows has been reported. After investigations, it appears this is due to some
printer information which are actually stored in the document itself when the
latter has been saved after it has been printed. If the printer doesn't exist
any more, the fall back is the default one\cite{writerPrinter1}. The print
shorcut might use the default WIndows printer\cite{writerPrinter2}. But the
user is able to choose if he wants to load the printer settings from the
document\cite{writerPrinter3}.

A new default font for LibreOffice for Arabic?\cite{arabicFont}

Implement regex for the REPLACE function in Calc to be OpenFormula compliant.
(this could be proposed as an easy hack)\cite{replaceRegexOpenFormula}.



\section{GSoC.}

Refactor god objects (Valentin): weekly report $\rightarrow$ pushed to master
and IDocumentChartDataProviderAccess has been nearly refactored.

Improve Usability of Personas (Rachit Gupta): the user can select the theme he
wants from the gallery and can be applied to the UI.  Needs to be done: run the
application of the theme in a separate thread to avoid the UI from hanging
+ little changes to the UI.

Connection to SharePoint and Microsoft OneDrive (Mihai Varga): all OneDrive is
serialized, a file and folder UI representation still needs to be writen. Now
writing unit tests.

Calc / Impress tiled rendering support (Andrzej Hunt): renamed to
LibreOfficeKit. For more details about the current implementation, please read
\cite{tiledRenderingArticle}.

As a reminder, tiled rendering is the ability to be able to paint (render)
a specific part of a document (a tile) to anywhere. This is really useful to
plug LibreOffice into various new scenarios where the traditional LO UI cannot
be used (mobile devices,\ldots).



\section{Hackfests and conferences.}

After the hackfest from Paris (Montreuil) hold on June 27-28th (thanks
Charles-H Schultz and Sophie Gautier), Toulouse has been confirmed to hold
a hackfest on November 15-16th\cite{hackfestToulouse} (thanks Arnaud Versini).
More to come at \cite{hackfestToulouseWiki}.

In 2015 linux.conf.au will be held 12-16 January in Auckland, New Zealand, at
the University of Auckland Business School.  Call for Proposals opened 9 June
2014, closes 13 July 2014\cite{linuxConfAuckland1}. There will not have
a LibreOffice booth during the conference itself, only on the open
day\cite{linuxConfAuckland2}.



\begin{thebibliography}{30}

\bibitem{wgetGithub}
\url{https://github.com/wget/lown}

\bibitem{gladeUi1}
\url{https://gerrit.libreoffice.org/9699}

\bibitem{gladeUi2}
\url{https://gerrit.libreoffice.org/9680}

\bibitem{gladeUi3}
\url{https://gerrit.libreoffice.org/9726}

\bibitem{sifrGithub1}
\url{http://listarchives.libreoffice.org/global/design/msg06641.html}

\bibitem{startCenterHidePreview1}
\url{https://bugs.freedesktop.org/show_bug.cgi?id=74834}

\bibitem{startCenterHidePreview2}
\url{http://listarchives.libreoffice.org/global/design/msg06640.html}

\bibitem{npapiMozilla}
\url{https://blog.mozilla.org/security/2014/02/28/update-on-plugin-activation/}

\bibitem{npapiLibONotWorking}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061635.html}

\bibitem{removeNpapiPatch}
\url{https://gerrit.libreoffice.org/#/c/9644/}

\bibitem{copyPasteDetectorBug}
\url{https://bugs.freedesktop.org/show_bug.cgi?id=39593}

\bibitem{copyPasteDetectorGithub}
\url{https://github.com/petersenna/ccfinderx-core}

\bibitem{access2BaseDefinition}
\url{http://www.access2base.com/access2base.html}

\bibitem{access2BaseOverridePatch}
\url{https://gerrit.libreoffice.org/#/c/9756/}

\bibitem{coreOverrideArguments}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061795.html}

\bibitem{extensionsCompatibility}
\url{http://listarchives.libreoffice.org/global/users/msg39416.html}

\bibitem{winBuildDependencies}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061742.html}

\bibitem{winBuildDependenciesTool}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061748.html}

\bibitem{winBuildDependenciesInstall}
\url{https://bugs.freedesktop.org/show_bug.cgi?id=32426}

\bibitem{winBuildGmakeReplaceGnuWin32}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061758.html}

\bibitem{winBuildCygwinMsysGnuWin32Comparison}
\url{http://stackoverflow.com/a/10712976}

\bibitem{localizedBuild}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061762.html}

\bibitem{freeBsdBuild1}
\url{https://gerrit.libreoffice.org/9714}

\bibitem{freeBsdBuild2}
\url{https://gerrit.libreoffice.org/9722}

\bibitem{memcheckTests}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061717.html}

\bibitem{sqlConnectorOsx1}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061700.html}

\bibitem{sqlConnectorOsx2}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061708.html}

\bibitem{writerPrinter1}
\url{http://listarchives.libreoffice.org/global/users/msg39393.html}

\bibitem{writerPrinter2}
\url{http://listarchives.libreoffice.org/global/users/msg39395.html}

\bibitem{writerPrinter3}
\url{http://listarchives.libreoffice.org/global/users/msg39396.html}

\bibitem{arabicFont}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061751.html}

\bibitem{replaceRegexOpenFormula}
\url{http://listarchives.libreoffice.org/global/users/msg39399.html}

\bibitem{hackfestToulouse}
\url{http://lists.freedesktop.org/archives/libreoffice/2014-June/061694.html}

\bibitem{hackfestToulouseWiki}
\url{https://wiki.documentfoundation.org/Hackfest/Toulouse2014}

\bibitem{linuxConfAuckland1}
\url{http://listarchives.libreoffice.org/global/marketing/msg16979.html}

\bibitem{linuxConfAuckland2}
\url{http://listarchives.libreoffice.org/global/marketing/msg16979.html}

\bibitem{opensslPatches}
\url{https://gerrit.libreoffice.org/#/c/9713/}

\bibitem{tiledRenderingArticle}
\url{http://www.ahunt.org/2014/05/tiled-rendering-with-libreofficekit/}

\end{thebibliography}

\end{document}
